Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-10 write-ups/PLDI/paper.tex
Add paragraph on adaptation of work stealing

@@

    level++;
    swap(in_nodes, out_nodes);
    reset(out_nodes);
    global_barrier();
  }
}
\end{lstlisting}
\caption{An OpenCL graph traversal algorithm, using a global barrier for synchronisation}\label{fig:graphsearch}
\end{figure}

@@

Figure~\ref{fig:graphsearch} shows the
skeleton of a graph traversal algorithm.  \ADComment{Describe some
  aspects of the example, saying that the global barrier is not a
  primitive, and can be coded up using the memory model but is not
  safe due to the execution model.  Make it clear that the barrier uses busy-waiting.}

@@

>>>>

(*@\label{line:graph:iterate}@*)  while(in_nodes.size > 0) {

@@

(*@\label{line:graph:swap}@*)    swap(in_nodes, out_nodes);
(*@\label{line:graph:gb1}@*)    global_barrier();
(*@\label{line:graph:reset}@*)    reset(out_nodes);
    level++;
(*@\label{line:graph:gb2}@*)    global_barrier();
  }
}
\end{lstlisting}
\caption{An OpenCL graph traversal algorithm, using a global barrier for synchronisation}\label{fig:graphsearch}
\end{figure}

@@

\ADComment{Describe some
  aspects of the example, saying that the global barrier is not a
  primitive, and can be coded up using the memory model but is not
  safe due to the execution model.  Make it clear that the barrier
  uses busy-waiting.}

\paragraph{Graph traversal} Figure~\ref{fig:graphsearch} shows the
skeleton of a frontier based graph traversal algorithm (this approach
has been shown to be efficient on GPUs in~\ref{...}). The kernel is
given three arguments in global memory: a graph structure, and two
arrays that can contain graph nodes. Initially, {\tt n0} contains the
starting nodes to process. Threads keep three local variables: {\tt
  level} which indicates the frontier level which is currently being
processed, {\tt in\_nodes} and {\tt out\_nodes} which point to
distinct node arrays. {\tt in\_nodes} contains the nodes to be
processed for the current frontier, while {\tt out\_nodes} contains
the nodes to be processed during the next frontier.

The application iterates as long as the current frontier contains
nodes to process (line~\ref{line:graph:iterate}). At each frontier,
the nodes to be processed are evenly distributed between the executing
threads. This is done through \emph{stride} based processing. A thread
retrieves its global thread id and the total number of threads
(obtained through the {\tt get\_global\_id()} and {\tt
  get\_global\_size()} OpenCL primitives). A thread then processes
each node associated with id (i.e.\ indices with its thread id value
added to any integer multiple of the total number of threads up to the
list size). A thread processes a node via the {\tt process\_node}
function, which may (a) push nodes to process in the next frontier to
the {\tt out\_nodes} list and (b) use the frontier level in
computation. After processing the frontier, the threads swap their
local node array pointers (line~\ref{line:graph:swap}), i.e.\ the {\tt
  out\_nodes} of the current frontier becomes the input of the next
frontier.

At this point, the GPU threads must wait for all other threads to
finishing processing the frontier before they can continue. To achieve
this, we use a global barrier construct
(line~\ref{line:graph:gb1}). After all threads reach this point, the
output node array is reset (line~\ref{line:graph:reset}) and the level
is incremented. Again, threads must wait until the output node is
reset at a global barrier (line~\ref{line:graph:gb2}). At this point
threads can continue to the next frontier.

The global barrier used in this application is not provided as a GPU
primitive, though many previous works have shown that such a global
barrier can be implemented~\cite{...}. These implementations use well
defined atomic memory operations to ensure memory consistency and use
spinning to ensure threads wait at the barrier for the remaining
threads to reach the barrier. Even though such barriers appear to run
reliably on current GPUs, they violate GPU language specifications,
which do not provide fair scheduling of workgroups. These barriers
require fair workgroup scheduling in order to function correctly. That
is, \emph{all} workgroups must be fairly scheduled in order to reach
the barrier, otherwise, the workgroups waiting at the barrier will not
make progress.

@@

For the work stealing example, there is no state to transmit since a
computation is entirely parametrized by a task, which is retrieved from
a task pool located in global memory. Care must be taken on the dynamic
change of the number of workgroups though: we cannot assume that the
total number of task pools is equal to $\getnumgroups$. Thus, the number
of pools must be given as an argument, and the $\mathsf{pool}$ argument
must be obtained from the group id modulo the total number of
pools. This total number of pools must also be passed to
$\mathsf{pop\_or\_steal}$, since when a workgroup wants to steal it must
access other pool indexes, which are bounded by the total number of
pools.
